Which PMV model should you use? Comparison of the ISO 7730:2005 and ASHRAE 55:2023 PMV models

### 2024-08-29 Manuscript/src/sections/introduction.tex
feat: updated the results using v_r instead of v

\section{Introduction}\label{sec:introduction}
\mycite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
The \ac{pmv} is an index that aims to predict the mean value of the thermal sensation votes (self-reported perceptions) of a large group of people on a sensation scale expressed from \num{-3} to \num{3} corresponding to the categories ``cold,'' ``cool,'' ``slightly cool,'' ``neutral,'' ``slightly warm,'' ``warm,'' and ``hot.''~\cite{iso7730, ashrae552023}.
The \ac{pmv} model has been widely used worldwide by researchers and practitioners.
It is the most extensively used thermal comfort index for estimating thermal sensation.
A search within the article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately \num{1600} documents.
Out of those \num{1082} are scientific peer-reviewed articles published in the research area of engineering, environmental science, social sciences, and energy.
This highlights this model's extensive adoption and use among the scientific community.
To this date, the \ac{pmv} remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988, tartarini_prediction_2023}.

To overcome the accuracy concerns, in particular when high air speed is used to cool people, the \gls{55} Standard uses a modified version of the original model, the \ac{pmv-ce}~\cite{ashrae552023}.
The best source to see how \ac{pmv-ce} works is the Standard itself, but the Standard does not explain why some choices were implemented.
A partial justification of the model is described in \mycite{arens_moving_2009} and \mycite{yang_cooling_2015}.
The process used to calculate the \ac{pmv-ce} in the \gls{55} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
In summary, when the \ac{v} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
The value of \ac{ce} is calculated using the \ac{set} equation.
The \ac{ce} is then subtracted from both the \ac{tdb} and \ac{tr}.
The resulting values become the new inputs in the \ac{pmv} model.
Since the \ac{ce} accounts for  convective heat losses from the person to its environment the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
The other three input parameters (\ac{clo}, \ac{met}, and \ac{rh}) remain unchanged.
Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
To quantify by how much the two models differ, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) for different values of \ac{v} using the \ac{pmv} and \ac{pmv-ce} models and we plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
%The rationale for this approach is that the original formulation of the \ac{pmv} model cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v} is higher than \qty{0.1}{\m\per\s}.
\begin{figure}[!htb]
    \centering
    \includegraphics[width=1\textwidth]{figures/pmv_comfort_regions}
    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models.
    The chart shows how the comfort regions change as a function of \ac{v}.
    We assumed $M$~=~1.2 met, $i_{cl}$~=~0.5 clo.
    \label{fig:comfort_regios_pmv_pmvce}}
\end{figure}

However, this method has limitations.
First, it should be noted that over the years, the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
Second, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the two-node model.
Third, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.
Finally, the \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} since the \ac{ce} already accounts for the convective heat losses, the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} normalizes them.
We could not find a rationale for this choice in the available literature.

\begin{figure}[!htb]
    \centering
    \begin{tikzpicture}[node distance=1.2cm]
        \node (start) [startstop] {\ac{pmv} calculation as per \gls{55}};
        \node (dec1) [decision, below of=start, yshift=-.25cm] {\ac{met} $> \qty{1}{met}$ };
        \node (pro1a) [process, right of=dec1, xshift=3.5cm] {\ac{v} = \ac{v}$+ 0.3 ($\ac{met}$-1)$};
        \node (dec2) [decision, below of=dec1, yshift=-0.75cm] {\ac{met} $> \qty{1.2}{met}$ };
        \node (pro2a) [process, right of=dec2, xshift=3.5cm] {\ac{clo} = \ac{clo}$(0.6 + 0.4/$\ac{met}$)$};
        \node (dec3) [decision, below of=dec2, yshift=-0.75cm] {\ac{v} $> \qty{0.1}{\m\per\s}$ };
        \node (end1) [startstop, below of=dec3, yshift=-.25cm, xshift=-4cm] {\ac{pmv}(\ac{tdb}, \ac{tr}, \ac{rh}, \ac{v}, \ac{met}, \ac{clo})};
        \node (pro3) [process, below of=dec3, yshift=-.25cm, xshift=3cm] {Calculate \ac{ce}};
        \node (end2) [startstop, below of=pro3, yshift=-.5cm] {\ac{pmv}(\ac{tdb} - \ac{ce}, \ac{tr} - \ac{ce}, \ac{rh}, \ac{v} = \qty{0.1}{\m\per\s}, \ac{met}, \ac{clo})};

        \draw [arrow] (start) -- (dec1);
        \draw [arrow] (dec1) -- node[above,pos=0.3] {Yes} (pro1a);
        \draw [arrow] (dec1) -- node[right,pos=0.3] {No} (dec2);
        \draw [arrow] (pro1a) -- ($(pro1a.south)+(0,-0.5)$) -| (dec2);
        \draw [arrow] (dec2) -- node[above,pos=0.3] {Yes} (pro2a);
        \draw [arrow] (dec2) -- node[right,pos=0.3] {No} (dec3);
        \draw [arrow] (pro2a) -- ($(pro2a.south)+(0,-0.5)$) -| (dec3);
        \draw [arrow] (dec3) -- ($(dec3.west)$) -| node[above,pos=0.3] {No} (end1);
        \draw [arrow] (dec3) -- ($(dec3.east)$) -| node[above,pos=0.3] {Yes} (pro3);
        \draw [arrow] (pro3) -- (end2);
    \end{tikzpicture}
    \caption{Flowchart depicting the steps for the calculation of the PMV following the \gls{55} Standard.}
    \label{fig:flowchart_pmv_ce}
\end{figure}

@@

Several other formulations of the \ac{pmv} model have been proposed to complicate this issue further.
Among the most notable formulations there are the: \ac{pmvs}~\cite{GaggeSET}, \ac{pmvg}~\cite{GaggeSET}, \ac{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
\mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describes some of the above-mentioned models.

@@

\subsection{Aim and Objectives}\label{subsec:aim-and-objectives}
In this paper, we aim to compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} Standards, respectively.
We used the self-reported \ac{tsv} recorded in the \acf{db2}.
We aim to determine which \ac{pmv} model is most accurate and which are its applicability limits.
This will allow \ac{pmv} users - engineers, architects, and educators - to make an informed decision on which model they should use and will help policymakers to select the most appropriate model to be used in Standards.

>>>>

\section{Introduction}\label{sec:introduction}
\mycite{Fanger1970} developed the \ac{pmv} model, which is now incorporated into the \gls{7730} Standard~\cite{iso7730} in its original form.
The \ac{pmv} is an index that aims to predict the mean value of the thermal sensation votes (self-reported perceptions) of a large group of people on a sensation scale expressed from \num{-3} to \num{3} corresponding to the categories ``cold,'' ``cool,'' ``slightly cool,'' ``neutral,'' ``slightly warm,'' ``warm,'' and ``hot.''~\cite{iso7730, ashrae552023}.
The \ac{pmv} model has been widely used worldwide by researchers and practitioners.
It is the most extensively used thermal comfort index for estimating thermal sensation.
A search within the article title, abstract, and keyword on Scopus with the search term ``predicted mean vote'' returns approximately \num{2050} documents.
Out of those \num{1200} are scientific peer-reviewed articles published in the research area of engineering, environmental science, social sciences, and energy.
This highlights this model's extensive adoption and use among the scientific community.
To this date, the \ac{pmv} remains to be the most utilized thermal comfort model even though several studies have highlighted that the \ac{pmv} has low accuracy in predicting thermal sensation votes~\cite{Cheung2019, Yao2022, kim2019thermal, tartarini2018thermal, Humphreys2002, doherty_evaluation_1988, tartarini_prediction_2023}.

To overcome the accuracy concerns, in particular when high air speed is used to cool people, the \gls{55} Standard uses a modified version of the original model, the \ac{pmv-ce}~\cite{ashrae552023}.
The best source to see how \ac{pmv-ce} works is the Standard itself, but the Standard does not explain why some choices were implemented.
A partial justification of the model is described in \mycite{arens_moving_2009} and \mycite{yang_cooling_2015}.
The process used to calculate the \ac{pmv-ce} in the \gls{55} is summarized in the flowchart depicted in Figure~\ref{fig:flowchart_pmv_ce}.
\begin{figure}[!htb]
    \centering
    \begin{tikzpicture}[node distance=1cm and 3cm, transform shape]
        \node (start) [startstop] {\acs{pmv} calculation as per \gls{55}};
        \node (dec1) [decision, below of=start, yshift=-.25cm] {\acs{met} $> \qty{1}{met}$ };
        \node (pro1a) [process, right of=dec1, xshift=3.5cm] {\acs{v} = \acs{v}$+ 0.3 ($\acs{met}$-1)$};
        \node (dec2) [decision, below of=dec1, yshift=-0.75cm] {\acs{met} $> \qty{1.2}{met}$ };
        \node (pro2a) [process, right of=dec2, xshift=3.5cm] {\acs{clo} = \acs{clo}$(0.6 + 0.4/$\acs{met}$)$};
        \node (dec3) [decision, below of=dec2, yshift=-0.75cm] {\acs{v} $> \qty{0.1}{\m\per\s}$ };
        \node (end1) [startstop, below of=dec3, yshift=-.25cm, xshift=-4cm] {\acs{pmv}(\acs{tdb}, \acs{tr}, \acs{rh}, \acs{v}, \acs{met}, \acs{clo})};
        \node (pro3) [process, below of=dec3, yshift=-.25cm, xshift=3cm] {Calculate \acs{ce}};
        \node (end2) [startstop, below of=pro3, yshift=-.5cm] {\acs{pmv}(\acs{tdb} - \acs{ce}, \acs{tr} - \acs{ce}, \acs{rh}, \acs{v} = \qty{0.1}{\m\per\s}, \acs{met}, \acs{clo})};

        \draw [arrow] (start) -- (dec1);
        \draw [arrow] (dec1) -- node[above,pos=0.3] {Yes} (pro1a);
        \draw [arrow] (dec1) -- node[right,pos=0.3] {No} (dec2);
        \draw [arrow] (pro1a) -- ($(pro1a.south)+(0,-0.5)$) -| (dec2);
        \draw [arrow] (dec2) -- node[above,pos=0.3] {Yes} (pro2a);
        \draw [arrow] (dec2) -- node[right,pos=0.3] {No} (dec3);
        \draw [arrow] (pro2a) -- ($(pro2a.south)+(0,-0.5)$) -| (dec3);
        \draw [arrow] (dec3) -- ($(dec3.west)$) -| node[above,pos=0.3] {No} (end1);
        \draw [arrow] (dec3) -- ($(dec3.east)$) -| node[above,pos=0.3] {Yes} (pro3);
        \draw [arrow] (pro3) -- (end2);
    \end{tikzpicture}
    \caption{Flowchart depicting the steps for the calculation of the PMV following the \gls{55} Standard.}
    \label{fig:flowchart_pmv_ce}
\end{figure}
In summary, when the \ac{v} exceeds \qty{0.1}{\m\per\s} the \gls{55} prescribes the use of the \ac{ce}.
The reader should note that \ac{v} is not the measured value.
The value of \ac{v} measured by the sensors, should be adjusted based on the \ac{met} of the occupant if \ac{met}~$>$~\qty{1}{met} using this equation \ac{v} = \ac{v}$+ 0.3 ($\ac{met}$-1)$.
The value of \ac{ce} is calculated using the \ac{set} equation.
The \ac{ce} is then subtracted from both the \ac{tdb} and \ac{tr}.
The resulting values become the new inputs in the \ac{pmv} model.
Since the \ac{ce} accounts for  convective heat losses from the person to its environment the value of \ac{v}, used to calculate the \ac{pmv}, is set to \qty{0.1}{\m\per\s}.
The other three input parameters (\ac{clo}, \ac{rh}, \ac{met}) remain unchanged.
Consequently, for a given thermal environment, results of the two \ac{pmv} formulations differ only when the value of \ac{v} is higher than \qty{0.1}{\m\per\s}.
To show to the reader by how much the output of the two models differ, we calculated the comfort regions ($\mid$\ac{pmv}$\mid \leq 0.5$) for different values of \ac{v} using the \ac{pmv} and \ac{pmv-ce} models assuming \ac{met}~=~\qty{1.2}{met} and \ac{clo}~=~\qty{0.5}{clo}.
We plotted the results in Figure~\ref{fig:comfort_regios_pmv_pmvce}.
\begin{figure}[!htb]
    \centering
    \includegraphics[width=1\textwidth]{figures/pmv_comfort_regions}
    \caption{Comfort regions ($|$\ac{pmv}$|$~$\leq$~\num{0.5}) calculated using \ac{pmv} and \ac{pmv-ce} models.
    The chart shows how the comfort regions change as a function of \ac{v}.
    We assumed $M$~=~1.2 met, $i_{cl}$~=~0.5 clo.
    \label{fig:comfort_regios_pmv_pmvce}}
\end{figure}
The results show that for \ac{v}~=~\qty{0.4}{\m\per\s} the comfort regions estimated using the \ac{pmv-ce} is wider than the one calculated using the \ac{pmv} and shifted towards the warmer temperatures.

The rationale for using the \ac{pmv-ce} model is that the original \ac{pmv} formulation cannot accurately estimate latent and sensible heat losses from the skin to the environment when \ac{v}~$>$~\qty{0.1}{\m\per\s}.
However, this method has its own limitations.
First, it should be noted that over the years, the critical value of \ac{v} has been changed, from \qty{0.15}{\m\per\s} in \mycite{Schiavon2014} to \qty{0.2}{\m\per\s} in the ASHRAE 55--2017~\cite{ASHRAE552017, arens_moving_2009} to \qty{0.1}{\m\per\s} in \gls{55}~\cite{ashrae552023}.
Second, calculating the \ac{pmv-ce} is more computationally intensive and complex since it requires the user to solve two heat balance equations, the \ac{pmv} and the two-node model.
Third, we could not find any peer-reviewed scientific publication that quantified the accuracy improvements of this model over the original \ac{pmv}.
Finally, the \ac{set} temperature which is used to calculate the \ac{ce}, is defined as the hypothetical isothermal environment at a \ac{rh} of \qty{50}{\percent}, \ac{v}~$\leq$~\qty{0.1}{\m\per\s}, and \ac{tr}~=~\ac{tdb} in which the total heat loss from the skin of an imaginary occupant wearing clothing, standardized for the activity concerned, is the same as that from a person in the actual environment with actual clothing and activity level~\cite{ashrae552023}.
However, while \ac{v}, used to calculate the \ac{pmv-ce}, is set to \qty{0.1}{\m\per\s} since the \ac{ce} already accounts for the convective heat losses, the value of \ac{rh} and \ac{clo} are not adjusted to compensate for the fact that \ac{set} normalizes them.
We could not find a rationale for this choice in the available literature.

Several other formulations of the \ac{pmv} model have been proposed to complicate this issue further.
Among the most notable formulations there are the: \ac{pmvs}~\cite{GaggeSET}, \ac{pmvg}~\cite{GaggeSET}, \ac{epmv}~\cite{Toftum2002}, and \ac{athb}~\cite{Schweiker2022}.
\mycite{Yao2022} provides a comprehensive review of different thermal comfort models and compares and describes some of the above-mentioned models.
Here, we decided to focus on the \ac{pmv} and \ac{pmv-ce} models since they are the most widely used thermal comfort models worldwide and are incorporated in the \gls{7730} and \gls{55} Standards.

@@

@@

\subsection{Aim and Objectives}\label{subsec:aim-and-objectives}
In this paper, we compare the accuracy of the \ac{pmv} and \ac{pmv-ce} models used in the \gls{7730} and \gls{55} Standards, respectively.
The paper objective is to quantitatively answer the following questions:
which model should you use?
Which model is more accurate?
We used the self-reported \ac{tsv} recorded in the \acf{db2}.
We aim to determine which \ac{pmv} model is most accurate and which are its applicability limits.
This will allow \ac{pmv} users - engineers, architects, and educators - to make an informed decision on which model they should use and will help policymakers to select the most appropriate model to be used in Standards.

### 2024-08-29 Manuscript/src/sections/methodology.tex
feat: wrote code to analyse data from studies with three air speed measurements
feat: updated the results using v_r instead of v

\section{Methodology}\label{sec:methodology}
The \ac{db2} has \var{entries_db_all} independent recordings from various field investigations throughout the world, and it is the largest of its kind~\cite{FoldvaryLicina2018, db2dryad}.
\Ac{tsv} were collected using a right-here-right-now survey and environmental parameters were measured and logged in the proximity of participants when they completed the survey.
The \ac{db2} only contains data collected by researchers in cross-sectional field studies that have been published in peer-reviewed journals.
 We compared the results obtained by the \ac{pmv} and the \ac{pmv-ce} with the self-reported \ac{tsv}.
Results were calculated using the function `pythermalcomfort.models.pmv\_ppd' included in pythermalcomfort v2.8.1 a Python package for thermal comfort calculations~\cite{Tartarini2020a}.

\subsection{Data Preparation and Cleaning}\label{subsec:data-processing-and-cleaning}
Not all researchers measured and logged all six parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv}.
This filtered out approximately \var{entries_db_valid} of the data.
We decided to keep entries that did not have \ac{tr}.
However, since \ac{tr} is a required input in the \ac{pmv} model, it was calculated using the following equation: $2 t_{o} – t_{db}$.
If the \ac{to} was not available for that entry we assumed \ac{tr} = \ac{tdb}.
We are aware that these assumptions may introduce an error, but removing samples without \ac{tr} would cause an additional \var{entries_db_valid_no_tr} drop in the data used to compare the models.
There is also evidence that in many conditions is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.

Both the \gls{55} and \gls{7730} specify a set of applicability limits, we hence filtered out those data points that did not meet the inclusion criteria of both Standards.
The rationale is that the models' accuracy should only be tested within their applicability ranges.
The limits we used are as follows:
\num{10}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius},
\num{10}~$\leq$~\ac{tr}~$\leq$~\qty{40}{\celsius},
\num{0}~$\leq$~\ac{v}~$\leq$~\qty{1}{\m\per\s},
\num{0}~$\leq$~\ac{clo}~$\leq$~\qty{1.5}{clo},
\num{0}~$\leq$~water vapour partial pressure~$\leq$~\qty{2000}{\pascal},
and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met}.
We then calculated the adjusted total clothing insulation and relative airspeed as required by both Standards.
We used these inputs to calculate the \ac{pmv} and \ac{pmv-ce} values.

@@

In addition, it should also be noted that while all the data included in the \ac{db2} is published in peer-reviewed papers, not all entries were collected using the same methodology.
For example, while both thermal comfort Standards recommend measuring \ac{v} at three different heights, this was not always done.
This could introduce a bias in the data.
Since the two \ac{pmv} formulations only differ when \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
\begin{enumerate}
    \item all the data points,
    \item all the data points with \ac{v}~$\geq$~\qty{0.2}{\m\per\s},
    \item only the data points with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
    These results are reported in the supplementary material.
\end{enumerate}
We believe that this approach will help us understand the impact of the data quality on the results of the models.
In particular, by only including those studies that measured \ac{v} at three different heights, we can reduce the bias introduced by the lack of adherence to the measurement protocol recommended by the Standards.

We used Python to analyse and visualise the results.
We are committed to reproducible research hence we have shared the source code and the dataset we used publicly at this URL: \url{https://github.com/FedericoTartarini/paper-pmv-comparison} so other users can test different assumptions.

>>>>

\section{Methodology}\label{sec:methodology}
The \ac{db2} has \var{entries_db_all} independent recordings from various field investigations throughout the world, and it is the largest of its kind~\cite{FoldvaryLicina2018, db2dryad}.
\Ac{tsv} were collected using a right-here-right-now survey and environmental parameters were measured and logged in the proximity of participants when they completed the survey.
The \ac{db2} only contains data collected by researchers in cross-sectional field studies that have been published in peer-reviewed journals.
 We compared the results obtained by the \ac{pmv} and the \ac{pmv-ce} with the self-reported \ac{tsv}.

\subsection{Data Preparation and Cleaning}\label{subsec:data-processing-and-cleaning}
Not all researchers measured and logged all six parameters needed to calculate the \ac{pmv} and \ac{pmv-ce}.
We, therefore, decided to remove all entries that did not have any of the following variables: \ac{tdb}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv}.
This filtered out approximately \var{entries_db_valid} of the data.
We decided to keep entries that did not have \ac{tr} and we planned to calculate it using the \ac{tdb} and \ac{to} values using the following equation: $2 t_{o} – t_{db}$.
If the \ac{to} was not available for that entry we assumed \ac{tr} = \ac{tdb}.
However, all the entries that had \ac{tdb}, \ac{rh}, \ac{v}, \ac{clo}, \ac{met}, and \ac{tsv} also had \ac{tr}.
Consequently, we did not need to make any assumptions about the \ac{tr} values.
In our analysis, we used the value \ac{v} which is adjusted as a function of \ac{met} and not the \ac{v} measured by the researchers.
This is because both \ac{pmv} formulations use the \ac{v} adjusted as a function of \ac{met} to calculate the \ac{pmv} value.
%We are aware that these assumptions may introduce an error, but removing samples without \ac{tr} would cause an additional \var{entries_db_valid_no_tr} drop in the data used to compare the models.
%There is also evidence that in many conditions is possible to assume \ac{tr} equal to \ac{tdb} without committing a major error~\cite{Dawe2020}.

Both the \gls{55} and \gls{7730} specify a set of applicability limits, we hence filtered out those data points that did not meet the inclusion criteria of both Standards.
The rationale is that the models' accuracy should only be tested within their applicability ranges.
The limits we used are as follows:
\num{10}~$\leq$~\ac{tdb}~$\leq$~\qty{30}{\celsius},
\num{10}~$\leq$~\ac{tr}~$\leq$~\qty{40}{\celsius},
\num{0}~$\leq$~\ac{v}~$\leq$~\qty{1}{\m\per\s},
\num{0}~$\leq$~\ac{clo}~$\leq$~\qty{1.5}{clo},
\num{0}~$\leq$~water vapour partial pressure~$\leq$~\qty{2000}{\pascal},
and \num{1}~$\leq$~\ac{met}~$\leq$~\qty{4}{met}.
The adjusted total clothing insulation and relative airspeed, as required by both Standards, were used to calculate the \ac{pmv} and \ac{pmv-ce} values.

@@

In addition, it should also be noted that while all the data included in the \ac{db2} is published in peer-reviewed papers, not all entries were collected using the same methodology.
For example, while both thermal comfort Standards recommend measuring \ac{v} at three different heights, this was not always done.
This could introduce a bias in the measurements and consequently in the results of the models.
Moreover, since the two \ac{pmv} formulations only differ when \ac{v}~$\geq$~\qty{0.1}{\m\per\s}, we decided to report three sets of results:
\begin{enumerate}
    \item all the data points,
    \item all the data points with \ac{v}~$\geq$~\qty{0.2}{\m\per\s},
    \item only the data points with \ac{v}~$\geq$~\qty{0.2}{\m\per\s} and from those studies that measured \ac{v} at three different heights.
    These results are reported in \ref{sec:analysis-of-the-dataset-with-air-speed-measurements-at-three-heights}.
\end{enumerate}
We believe that comparing the results from these three datasets will help us understand the impact of the data quality on the results of the models.
In particular, by only including those studies that measured \ac{v} at three different heights, we can reduce the bias introduced by the lack of adherence to the measurement protocol recommended by the Standards.

We used Python to analyse and visualise the results.
The \ac{pmv} results were calculated using the function `pythermalcomfort.models.pmv\_ppd' included in pythermalcomfort v2.8.1 a Python package for thermal comfort calculations~\cite{Tartarini2020a}.
We are committed to reproducible research hence we have shared the source code and the dataset we used publicly at this URL: \url{https://github.com/FedericoTartarini/paper-pmv-comparison} so other users can test different assumptions.
